---
title: "p8105_hw1_rs4390"
author: Rae Spriggs
output: github_document
date: "2022-09-17"
---


# Problem 1


```{r load data}
data("penguins", package = "palmerpenguins")
```

```{r, echo = FALSE, message = FALSE}
 library(tidyverse)
```

## Dataset description
The penguins dataset contains `r nrow(penguins)` observations of individual penguins. There are `r ncol(penguins)` total variables that detail the penguins' physical dimensions, sex, year, species, and island. 

The penguins in the dataset are from 3 islands (Biscoe, Dream, and Torgersen), 3 years (2007, 2008, and 2009), and represent 3 species (Adelie, Chinstrap, and Gentoo). The other variables include: bill length, bill depth, flipper length, body mass, and sex. The mean flipper length of the dataset is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE))` mm. 

## Scatterplot 

```{r scatterplot, echo = FALSE, message=FALSE}
penguins %>%
  ggplot(aes(
    x = bill_length_mm, 
    y = flipper_length_mm, 
    color = species)) + 
  geom_point() + 
  xlab("Bill Length (mm)") + 
  ylab("Flipper Length (mm)") + 
  labs(color = "Penguin Species")

ggsave("hw1_penguins_scatterplot_rs4390.png")
```

# Problem 2 

## New data frame 
```{r data frame}

hw1_df = 
  hw1_df = tibble(
    norm_samp = rnorm(10), 
    norm_samp_pos = norm_samp > 0, 
    var_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
    var_factor = factor(c("male", "female", "nonbinary", "male", "male", "nonbinary", "male", "male", "female", "nonbinary")),
  )

x <- c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J")
vectorLength = length(x)
cat("Length of vector x is :", vectorLength)

y <- c("male", "female", "nonbinary", "male", "male", "nonbinary", "male", "male", "female", "nonbinary")
vectorLength = length(y)
cat("Length of vector y is :", vectorLength)
  
```

I created a data frame comprised of:

* a random sample of size 10 from a standard Normal distribution with `rnorm`
* a logical vector indicating whether elements of the sample are greater than 0 with `norm_samp_pos`
* a character vector of length 10 with `var_char`
* a factor vector of length 10, with 3 different factor “levels” with `var_factor`

## Taking mean of each variable

```{r}

```

